---
title: "Data Download and Cleaning"
author: "Subrata Paul"
date: "6/27/2020"
output: html_document
---
### Function to download the data directly fro website. 
It creates a directory by year and saves the data within the directory. 
```{r eval=F}
download_data = function(year, ext){
  data_names<-paste(c('UTAS','UAS','UTASS','UASS', 'TRICH','CHLMDA','DEMO','DIQ','HIV','ALQ', 'WHQ'), ext, sep = '_')
  #data_names<-c('UTASS_H','UASS_H', 'TRICH_H','CHLMDA_H','DEMO_H','DIQ_H','HIV_H','ALQ_H')
  system(paste0('mkdir /Users/spaul/Documents/Arsenic/arsenic/data/', year))
  setwd(paste0('/Users/spaul/Documents/Arsenic/arsenic/data/',year))
  system('rm *')
  system(paste(paste('wget https://wwwn.cdc.gov/Nchs/Nhanes/',year,'/',data_names,'.XPT ',sep=''),collapse  = ' \ '))
}


```

```{r results='hide', message=F, warning=F, error = F, collapse=TRUE, eval=F}
download_data('2015-2016', 'I')
download_data('2013-2014', 'H')
download_data('2011-2012','G')
download_data('2009-2010', 'F')
download_data('2007-2008', 'E')
download_data('2005-2006', 'D')

```


```{r}
create_result<-function(lmod){
  lmod$coefficients
  out<-data.frame(Coef = lmod$coefficients)
  out$p_values = summary(lmod)$coefficients[,4]
  out$OR = exp(coef(lmod))
  out = cbind(out, confint(lmod))
  return(out)
}
```


### Prepare covariates. Note that not all the covariates are used in the analysis e.g. diabetes. The following transformations are performed. 

* On 2015-2016 data HIV has four variables, HIV1, HIV2 and HIV confirmatory test. We used the confirmatory test data. 
* BMI is calculated as (https://www.cdc.gov/healthyweight/assessing/bmi/childrens_bmi/childrens_bmi_formula.html) $$BMI = 730\times \frac{wight(lb)}{(height(in))^2}$$
* Depending on variable values 7,8,77,99,7777,9999 are interpreted as missing
* PIR is converted to categorical variables with break points 1.3, and 3.5
* AGE is converted to categorical with break points 20, 40, and 65
* BMI is converted to categorical with break points  18.5, 24.9, 29.9
* EDU2 is used. Level 2 and 3 are merged to have the same level. 

Bellow is the function to perform the above tasks on each year dataset. 
```{r}
library(SASxport)
get_covar = function(year, ext){
  setwd(paste0('/Users/spaul/Documents/Arsenic/arsenic/data/',year))
  DEMO<-read.xport(paste0('DEMO_',ext,'.XPT'))
  
  names(DEMO)[grep('BORN',names(DEMO))]<-'DMDBORN'
  DEMO$DMDBORN[DEMO$DMDBORN%in%c(7,9,77,99)]<-NA
  DEMO$DMDBORN[DEMO$DMDBORN>1]<-2
  DIQ<-read.xport(paste0('DIQ_',ext,'.XPT'))
  HIV<-read.xport(paste0('HIV_',ext,'.XPT'))
  if(year == '2015-2016') names(HIV)[5] = 'LBDHI'
  ALQ<-read.xport(paste0('ALQ_',ext,'.XPT'))
  WHQ = read.xport(paste0('WHQ_',ext,'.XPT'))
  WHQ$WHD010[WHQ$WHD010%in%c(7777,9999)]<-NA
  WHQ$WHD020[WHQ$WHD010%in%c(7777,9999)]<-NA
  WHQ$BMI = 730*as.numeric(WHQ$WHD020) / as.numeric(WHQ$WHD010^2)
  
  
  covar <- merge(DEMO[,c('SEQN','RIAGENDR','RIDAGEYR','RIDRETH1','DMDBORN','DMDYRSUS','DMDEDUC2','DMDEDUC3', 'INDFMPIR')], DIQ[,c('SEQN','DIQ010')], all.x = T)
  covar <- merge(covar, HIV[,c('SEQN','LBDHI')], all.x = T)
  covar <- merge(covar, ALQ[,c('SEQN','ALQ101')], all.x = T)
  covar <- merge(covar, WHQ[,c('SEQN','BMI')], all.x = T)
  
  #covar$DMQADFC[covar$DMQADFC%in%c(7)]<-NA
  #covar$DMDBORN[covar$DMDBORN4==99]<-NA
  covar$DMDYRSUS[covar$DMDYRSUS%in%c(77,99)]<-NA
  covar$DMDEDUC2[covar$DMDEDUC2%in%c(7,9)]<-NA
  covar$DMDEDUC3[covar$DMDEDUC3%in%c(99)]<-NA
  #covar$INDHHIN2[covar$INDHHIN2S%in%c(77,99)]<-NA
  covar$DIQ010[covar$DIQ010%in%c(9)]<-NA
  covar$ALQ101[covar$ALQ101%in%c(9)]<-NA
  covar$YEAR = rep(year,nrow(covar))
  
  covar$SEQN <- as.integer(covar$SEQN)
  covar$LBDHI[covar$LBDHI==3]<-NA
  covar$LBDHI <- as.factor(covar$LBDHI)
  covar$RIAGENDR<-as.factor(covar$RIAGENDR)
  covar$DMDBORN<-as.factor(covar$DMDBORN)
  covar$DMDYRSUS <- as.numeric(covar$DMDYRSUS)
  
  covar$INDFMPIR = cut(covar$INDFMPIR, breaks = c(0,1.3, 3.5,Inf), labels = c(1,2,3))
  covar$RIDAGEYR = cut(covar$RIDAGEYR, breaks = c(0, 20, 40, 65, Inf), labels = c(1,2,3,4))
  covar$RIDRETH1 = as.factor(covar$RIDRETH1)
  levels(covar$RIDRETH1) <- c("1","1","2","3","4")
  
  covar$BMI = cut(covar$BMI, breaks = c(0,18.5, 24.9, 29.9, Inf), labels = c(1,2,3,4))
  covar$DMDEDUC2 <- as.factor(covar$DMDEDUC2)
  levels(covar$DMDEDUC2)<-c("1","2","2","3","4")
  covar$DMDEDUC3<-as.factor(covar$DMDEDUC3)
  covar$DIQ010 <- as.factor(covar$DIQ010)
  covar$ALQ101 <- as.factor(covar$ALQ101)
  return(covar)
}
```

### Collect covariates and combine them together

```{r}
covar = list()
covar$Y2015 = get_covar('2015-2016','I')
covar$Y2013 = get_covar('2013-2014','H')
covar$Y2011 = get_covar('2011-2012','G')
covar$Y2009 = get_covar('2009-2010','F')
covar$Y2007 = get_covar('2007-2008','E')
covar$Y2005 = get_covar('2005-2006','D')
covar$ALL = rbind(covar$Y2015, covar$Y2013, covar$Y2011, covar$Y2009, covar$Y2007, covar$Y2005)
```

#### Class of the covariates
```{r}
sapply(covar$ALL, class)
```

#### Summary Statistics of the covariates
```{r}
get_summary = function(x){
  if(class(x)=='numeric') return(summary(x))
  if(class(x)=='integer'){
    if(length(unique(x))<=17){
      return(table(x, useNA = 'ifany'))
    }else{
      return(summary(x))
    }
  }
  if(class(x)=='factor') return(table(x, useNA = 'ifany'))
}

sapply(covar$ALL, get_summary)
```

### Prepare the arsenic data
We used both the total and speciated arsenic from the special subsample (if available). special subsample were available from 2011. TUC1 and TUC2 are created as 

TUC1 = Arsenous + Arsenic + Dimethylarsinic + Monomethylacrsonic 
TUC2 = TOTAL - Arsenobetaine - Arsenocholine

```{r message=F, warning=F}
year = '2013-2014'
ext = "H"
get_arsenic_data = function(year, ext){
  setwd(paste0('/Users/spaul/Documents/Arsenic/arsenic/data/',year))
  if(file.exists(paste0('UTASS_',ext,'.XPT'))){
    total<-read.xport(paste0('UTASS_',ext,'.XPT'))
    esp<-read.xport(paste0('UASS_',ext,'.XPT'))
    arsenic = merge(total, esp[, names(esp)!="WTFSM"])
  }else if (file.exists(paste0('UASS_',ext,'.XPT'))){
    arsenic<-read.xport(paste0('UASS_',ext,'.XPT'))
  }else{
    arsenic<-read.xport(paste0('UAS_',ext,'.XPT'))
  }

    arsenic = arsenic[,- grep('LC',names(arsenic))]
    #if('URDUMMAL'%in%names(arsenic)) arsenic = subset(arsenic, select = -URDUMMAL)
    #if('URXUCR'%in%names(arsenic)) arsenic = subset(arsenic, select = -URXUCR)
    #if('URXUCR'%in%names(arsenic)) arsenic = subset(arsenic, select = -URXUCR)
    names(arsenic)[names(arsenic)=='WTSA2YR'] = 'WTFSM'
    arsenic = arsenic[,c("SEQN","WTFSM","URXUAS","URXUAS3","URXUAS5","URXUAB","URXUAC","URXUDMA","URXUMMA")]
    arsenic$TUC1 = arsenic$URXUAS3 + arsenic$URXUAS5 + arsenic$URXUMMA + arsenic$URXUDMA
    arsenic$TUC2 = arsenic$URXUAS - arsenic$URXUAB - arsenic$URXUAC
    arsenic$YEAR = rep(year, nrow(arsenic))
    
  return(arsenic)
}

arsenic<-list()
arsenic$Y2015 = get_arsenic_data('2015-2016', 'I')
arsenic$Y2013 = get_arsenic_data('2013-2014', 'H')
arsenic$Y2011 = get_arsenic_data('2011-2012', 'G')
arsenic$Y2009 = get_arsenic_data('2009-2010', 'F')
arsenic$Y2007 = get_arsenic_data('2007-2008', 'E')
arsenic$Y2005 = get_arsenic_data('2005-2006', 'D')
arsenic$ALL = rbind(arsenic$Y2015, arsenic$Y2013, arsenic$Y2011 , arsenic$Y2009, arsenic$Y2007, arsenic$Y2005)
```

## Prepare Dependent variables

```{r, message=F, warning=F}
get_chlmda_data= function(year, ext){
  setwd(paste0('/Users/spaul/Documents/Arsenic/arsenic/data/',year))
  chlmda = read.xport(paste0('CHLMDA_',ext,'.XPT'))
  chlmda = chlmda[,c("SEQN", "URXUCL")]
  chlmda$URXUCL = as.factor(chlmda$URXUCL - 1 )
  return(chlmda)
}

chlmda<-list()
chlmda$Y2015 = get_chlmda_data('2015-2016', 'I')
chlmda$Y2013 = get_chlmda_data('2013-2014', 'H')
chlmda$Y2011 = get_chlmda_data('2011-2012', 'G')
chlmda$Y2009 = get_chlmda_data('2009-2010', 'F')
chlmda$Y2007 = get_chlmda_data('2007-2008', 'E')
chlmda$Y2005 = get_chlmda_data('2005-2006', 'D')
chlmda$ALL = rbind(chlmda$Y2015, chlmda$Y2013, chlmda$Y2011 , chlmda$Y2009, chlmda$Y2007, chlmda$Y2005)
```

## Prepare regression data

```{r}
reg_data = merge(chlmda$ALL, arsenic$ALL)
reg_data = merge(reg_data, covar$ALL)
```
# Analysis on Chlamydia
## Speciated arsenic model

```{r, results='asis'}
library(xtable)
lmod_esp = glm(URXUCL ~ WTFSM + URXUAS3 + URXUAS5 + URXUAB + URXUAC + URXUDMA + URXUMMA + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_esp), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```
#### Results with OR and 95% confidence interval

```{r, results='asis', message=F, warning=F}
print(xtable(create_result(lmod_esp), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

## Total arsenic model

```{r, results='asis'}
lmod_total = glm(URXUCL ~ WTFSM + URXUAS + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_total), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')

```


#### Results with OR and 95% confidence interval

```{r, results='asis', message=F}
print(xtable(create_result(lmod_total), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

## TUC1 model 

```{r, results='asis'}
lmod_tuc1 = glm(URXUCL ~ WTFSM + TUC1 + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_tuc1), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

#### Results with OR and 95% confidence interval

```{r, results='asis', message=F}
print(xtable(create_result(lmod_tuc1), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```


## TUC2 model 

```{r, results='asis'}
lmod_tuc2 = glm(URXUCL ~ WTFSM + TUC2 + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_tuc2), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```
#### Results with OR and 95% confidence interval

```{r, results='asis', message=F}
print(xtable(create_result(lmod_tuc2), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```


## Prepare Dependent variables

```{r, message=F, warning=F}
get_trich_data= function(year, ext){
  setwd(paste0('/Users/spaul/Documents/Arsenic/arsenic/data/',year))
  trich = read.xport(paste0('TRICH_',ext,'.XPT'))
  trich = trich[,c("SEQN", "URXUTRI")]
  trich$URXUTRI = as.factor(trich$URXUTRI - 1 )
  return(trich)
}

trich<-list()
trich$Y2015 = get_trich_data('2015-2016', 'I')
trich$Y2013 = get_trich_data('2013-2014', 'H')
trich$ALL = rbind(trich$Y2015, trich$Y2013)
```



```{r}
reg_data = merge(trich$ALL, arsenic$ALL)
reg_data = merge(reg_data, covar$ALL)
```

# Analysis on Trichomonas
## Speciated arsenic model

```{r, results='asis'}
lmod_esp = glm(URXUTRI ~ WTFSM + URXUAS3 + URXUAS5 + URXUAB + URXUAC + URXUDMA + URXUMMA + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_esp), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

#### Results with OR and 95% confidence interval

```{r, results='asis', message=F, warning=F}
print(xtable(create_result(lmod_esp), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

## Total arsenic model

```{r, results='asis'}
lmod_total = glm(URXUTRI ~ WTFSM + URXUAS + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_total), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

#### Results with OR and 95% confidence interval

```{r, results='asis', message=F, warning=F}
print(xtable(create_result(lmod_total), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```



## TUC1 model 

```{r, results='asis'}
lmod_tuc1 = glm(URXUTRI ~ WTFSM + TUC1 + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_tuc1), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```
#### Results with OR and 95% confidence interval

```{r, results='asis', message=F, warning=F}
print(xtable(create_result(lmod_tuc1), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

## TUC2 model 

```{r, results='asis', message=F, warning=F}
lmod_tuc2 = glm(URXUTRI ~ WTFSM + TUC2 + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDBORN   + DMDEDUC2 + INDFMPIR  + ALQ101 + BMI , data = reg_data, family = binomial())
print(xtable(summary(lmod_tuc2), align = "lcccc"), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```

#### Results with OR and 95% confidence interval

```{r, results='asis', message=F, warning=F}
print(xtable(create_result(lmod_tuc2), align = 'lccccc'), type = 'html', 
      html.table.attributes = 'align="center", rules = "rows", width = 50%, frame="below"')
```


